Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2020-10-28 ms/sections/abstract.tex
backup

Animal signals evolve in an ecological context. Moreover, locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation. Dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers, albeit at relatively large geographical scales. At smaller scales, gene flow is likely to erase phenotypic differentiation between localities. Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, as this may give new insights into the local scale at which adaptation is possible. We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands. Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances. Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results. On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was falsified. We suggest, however, that selection could be involved in several ways, including sexual selection. Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

>>>>

Animal signals evolve in an ecological context. Locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation. Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales. Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, as this may give new insights into the local scale at which adaptation is possible. We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands. Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances. Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results. On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported. We suggest, however, that selection could be involved in several ways, including sexual selection. Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

### 2020-10-28 ms/sections/acknowledgments.tex
backup

Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, and the Cayman Islands Department of the Environment. The authors thank Sofia Prado-Irwin, Pavitra Muralidhar, Nicholas Herrmann, Richard E. Glor, Alberto R. Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions. Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

>>>>

Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, the Bahamas Ministry of Agriculture, and the Cayman Islands Department of the Environment. The authors thank Sofia Prado-Irwin, Pavitra Muralidhar, Nicholas Herrmann, Richard E. Glor, Alberto R. Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions. Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

### 2020-10-28 ms/sections/discussion.tex
backup

% There are inconsistent differences that are not explained by distance

%firef\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence. Simple models of genetic drift are ruled out and we don’t think it is plasticity. However, our primary test for adaptation—looking for parallel responses across islands—also was falsified. We suggest, however, that selection could be involved, and suggest several ways this might have happened.}

\paragraph{Dewlap coloration differs between habitat-types} We found that male dewlap coloration in \textit{A. sagrei} significantly varied between fine-scale habitat-types (beach scrub bush, primary coppice forest and mangrove forest) on five islands of the West Indies: Abaco, Bimini, Cayman Brac, Little Cayman and Long Island. However, the habitat-specific variation in dewlaps was not consistent between these islands. Although those results are consistent with selection acting at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or multiple colonization events. We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).

\paragraph{A role of neutral drift is unlikely} Differences in organismal traits between environments are not necessarily proof of adaptation or selection, and genetic drift may result in patterns similar to local adaptation \citep{Miles2019}. Nevertheless, two lines of evidence from our data suggest that this scenario may be implausible. First, we found little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report. We did detect a significant signal of isolation-by-distance on Eleuthera, but there were no differences in dewlap coloration between habitats on this island. Second, we detected differences between habitats at relatively small spatial scales, most of the time between sites 5-10km apart, sometimes a few hundred meters away, on Bimini for example. Such small-scale differences would be unlikely under strong gene flow \citep{Richardson2014}. Our study islands lack geographic barriers to the movement of \textit{A. sagrei}, which have been shown to be highly mobile \citep{Kamath2018}, implying widespread gene flow across sites and habitats. Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al. unpublished thesis), suggesting gene flow between habitats may be widespread. Our results align with previous documented cases of persistent dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions. \citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration in the face of gene flow between subspecies of \textit{A. distichus} across Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}. \citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A. apletophallus} from Panama. \citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A. roquet} on Martinique, and a convergent pattern was found in \textit{A. trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}. Finally, regionally-distinct body coloration, but not dewlap coloration, is present in \textit{A. conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.

% \paragraph{No conclusive evidence for local adaptation of dewlap coloration} Although phenotypic divergence at small spatial scales may be an indicator of local adaptation, the inconsistency of the between-habitat divergence patterns we observed across islands implies that neutral processes cannot be definitely ruled out \citep{Losos2011}. Conversely, the absence of parallel divergence across islands does not rule out local adaptation either \citep{Losos2011}. If selection has an effect, it may be dependent on aspects of the environment that are not encompassed by our broad habitat-type classification into coastal dry scrub, primary coppice forest and mangrove forest, and dewlap coloration may be influenced by a mosaic of interacting local environmental factors, which need not be the same across islands.\\

\paragraph{No conclusive evidence for adaptation of dewlap coloration to habitat type} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}. Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}. However, the inconsistent and idiosyncratic patterns we observed reject the test of convergence across islands, suggesting that dewlap color variation between habitats cannot be predicted by habitat identity alone in \textit{A. sagrei}. Other studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e. UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A. cristatellus} and \textit{A. cooki} from Puerto Rico \citep{Leal2002, Leal2004}. Although UV reflectance was commonly involved in between-habitat divergence in \textit{A. cooki} and \textit{A. cristatellus}, we found no such patterns in \textit{A. sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}. Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}. Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A. sagrei}. Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}. In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A. sagrei} on islands with more coexisting species of anoles. Therefore, if local adaptation occurs, it is more likely to involve components of the environment that do not encompass our broad habitat categories.

\paragraph{Sexual selection could be at play} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Andersson1994}. This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g. plasticity or random drift). Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario. However, \citet{Baeckens2018} found no link between \textit{A. sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos. Another form of sexual selection is the "good genes" model, where the cue under sexual selection is an indicator of individual quality (e.g. better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}. Several studies suggest that this possibility could be the case for anoles. For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A. brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition. \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A. sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health. Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}. However, one would expect good genes sexual selection to favor the same coloration across the archipelago, and therefore this mechanism is unlikely to have produced idiosyncratic patterns of within-island divergence.

% An alternative yet non-mutually exclusive explanation for the observed idiosyncratic patterns of within-island divergence is that of "Fisherian" sexual selection \citep{Andersson1994}.

% Under the Fisherian model of sexual selection, arbitrary female preferences (i.e. independent of the environment) for certain male ornaments may drive divergent evolution of male signals, such as dewlap coloration, if female preferences differ between localities for other reasons than environment-dependent perception abilities (a situation that could have arisen e.g. in an initial phase of genetic drift). Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this explanation. Therefore, arbitrary, habitat-independent female preferences could further explain the inconsistent patterns of divergence across islands. In contrast with this, however, \citet{Baeckens2018} found no link between \textit{A. sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.\\

% Alternatively, dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection, rather than a Fisherian process. In the good genes model, the cue under sexual selection is an indicator of individual quality (e.g. better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}. Several studies suggest that this possibility could be the case for anoles. For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A. brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition. \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A. sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health. Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.\\

\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}. The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}. However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A. sagrei} \citep{Steffen2010} nor in \textit{A. distichus} \citep{Ng2013}, which also has an orange-based dewlap. This makes a plastic response to differences in diet across habitats unlikely. Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}. The differences we observed could therefore be genetically based. This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A. sagrei}. % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore

% For these reasons, we consider implausible that dewlap coloration is a plastic trait in our study species.\\

% Too broad a statemtent to make
% Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.

% This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.

% Conclusion, and a word about speciation

\paragraph{Implications in the context of speciation} Local adaptation can be a precursor to ecological speciation, a process that may have given rise to the adaptive radiation of \textit{Anolis} lizards \citep{Harmon2003, Gavrilets2009}. Ecologically-mediated divergence of a sexual signal may be a potent path to the evolution of reproductive isolation through divergent sexual selection \citep{Reynolds2007, Servedio2011}. Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}. Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular has been capable of rapid macroevolution throughout its phylogenetic history \citep{Nicholson2007}. For example, \textit{A. conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}. We present evidence of multiple cases of within-island dewlap color divergence over small geographical scales in \textit{A. sagrei} across the West Indies. While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap could have enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.

>>>>

\paragraph{Dewlap coloration differs between habitats, but the differentiation is inconsistent with adaptive convergence} Dewlaps are highly conspicuous in anoles and used as communication signals in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013}. On five of the nine islands we sampled in the West Indies, we found that \textit{A. sagrei} male dewlap coloration significantly varied between individuals living in three different habitat types  (beach scrub bush, primary coppice forest and mangrove forest), even though those habitats were in close proximity to each other and the \textit{sagrei} populations are continuously distributed between the habitats.\\

One compelling indicator of adaptation by natural selection is the repetition of patterns of differentiation across localities \citep{Losos2009, Losos2011}. In the context of this study, had we found consistent patterns of dewlap differentiation between habitats, we would have concluded that there was likely an adaptive basis for this pattern. However, the inconsistent and idiosyncratic patterns of dewlap variation we observed contradict this prediction. What then explains the observed differences among the populations?\\

One explanation is that this pattern is the result of multiple island colonizations of phenotypically different individuals to these habitats. However, we reject this possibility because all of the island populations in this study are strictly monophyletic, reflecting a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}). Other drivers could be at work, including genetic drift, phenotypic plasticity, and divergent natural selection. Of these, we find divergent natural selection the most likely explanation, suggesting that future research with additional granularity on the drivers of dewlap coloration is needed.

\paragraph{A role of neutral drift is unlikely} Genetic drift may lead to patterns of trait divergence similar to local adaptation \citep{Miles2019}. We think this scenario is implausible, however, as we found significant differences in dewlap color over distances too small for genetic drift to have realistically counteracted the effect of homogenizing gene flow \citep{Richardson2014}. On Bimini, for example, we detected differences between sites only a few hundred meters away and \textit{A. sagrei}, have previously been shown to be highly mobile \citep{Kamath2018}. Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA (van de Schoot et al. unpublished thesis). This is further evidence of gene flow between habitats.\\

It is possible that complex drift scenarios other than simple distance-mediated, diffusion-like processes may be at play \citep{Miles2019}. For example,  we did detect a significant signal of isolation-by-distance on Eleuthera, without finding differences in dewlap coloration between habitats, suggesting a possible role of drift on this island. Nonetheless, by and large our results align with previously documented cases of highly localized  dewlap color divergence despite gene flow in multiple species of anoles. \citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration despite gene flow between subspecies of \textit{A. distichus} on Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}. \citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A. apletophallus} from Panama. \citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A. roquet} on Martinique, and a convergent pattern was found in \textit{A. trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}. Finally, regionally-distinct body coloration, but not dewlap coloration, is present in \textit{A. conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.

\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors. The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}. However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A. sagrei} \citep{Steffen2010} nor in \textit{A. distichus} \citep{Ng2013}, which also has an orange-based dewlap. (Dewlap color has also been shown to change in relation to parasite load \citep{Cook2013}, though parasites were not investigated in this study. -- to remove) These laboratory results  suggest a plastic response in dewlap color to differences in diet across habitats unlikely, suggesting the differences we observed could therefore be genetically based. This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A. sagrei}.

\paragraph{Sensory Drive - maybe drop?} One hypothesis that a relationship would exist between dewlap color and habitat stems from the idea that the communication signals evolve adaptively in response to light environment (the sensory drive hypothesis, \citealt{Endler1988, Endler1992, Endler1998}). However, we find this explanation unlikely because the differences we observed were both inconsistent among islands and inconsistent with predictions of the sensory drive hypothesis. Previous studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions, primarily through UV contrast (i.e. UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{ A. cristatellus} and \textit{A. cooki} from Puerto Rico \citep{Leal2002, Leal2004}. On the contrary, we found no apparent habitat-dependent maximization of UV-contrast, or just any contrast in \textit{ A. sagrei}. Instead, we found for example the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}. Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}. Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A. sagrei}.

\paragraph{Fisherian sexual selection - maybe drop?} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Lande1981, Andersson1994, Higashi1999}. This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g. plasticity or random drift). Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario. However, \citep{Baeckens2018} found no link between \textit{A. sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.

\paragraph{Natural selection may proceed differently across islands and habitats} A further potential explanation for the inconsistencies in dewlap coloration among habitats between islands is that the observed differences in dewlap color are the result of divergent natural selection. If that is the case, however, either the habitats we have identified are not as similar as we thought, such that the same habitat type on different islands favors different colors, or, alternatively, the selective pressures causing dewlap divergence are related to some environmental factor other than habitat type that we did not measure. The habitats on different islands may, for example, differ in densities of predators or congeners, which have been shown to affect among-island dewlap diversity \cite{Vanhooydonck2009, Baeckens2018}. In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A. sagrei} on islands with more coexisting species of anoles. Therefore, if local adaptation occurs, it may be more likely to involve components of the environment that do not encompass our broad habitat categories leading to island-specific differences between habitats rather than parallel differences expected if habitat identity was the primary selective pressure driving divergence.\\

One last possibility to explain idiosyncratic patterns of between-habitat differences we observe is that, if island populations are different in color on average, they may be experiencing different portions of a within-island fitness landscape whose shape changes depending on the starting phenotype. In this study, we indeed found that dewlap coloration differed much more between than within islands. So, a population may not respond to selection in the same manner depending on its starting conditions. This is true not only in terms of phenotype (average dewlap color differs among islands) but also genetic composition (e.g. independent colonizations may have important genetic founder effects, \citealt{Reynolds2020}), and so the selective pressures applying to populations with different starting phenotypes may drive these divergent outcomes. For example, \citet{Fleishman2020} found that yellow stimuli are less detectable than red stimuli in high-light environments while both colors are readily detectable in low-light environments. Consequently, the environmental pressures applying to yellower dewlaps may differ than those for redder dewlaps, as red dewlaps should be always equally or more detectable than yellow dewlaps across light environments, whereas yellow dewlaps could experience more directional selection towards more red when in a high-light environments.

\paragraph{Conclusion} We identified patterns of dewlap color differentiation in populations of \textit{A. sagrei} from different habitat-types on multiple small islands of the West Indies. However, our main hypothesis to explain our findings -- local adaptation and sensory drive -- received little support from our multiple-island dataset. We also found other mechanisms such as drift or plasticity unlikely candidates to explain the observed patterns. This suggests that combinations of multiple factors interacting in subtle ways, and not necessarily in parallel across islands, may have contributed to present-day dewlap color diversity.

### 2020-10-28 ms/sections/introduction.tex
backup

The staggering diversity of animal communication signals has long been of interest to evolutionary biologists. Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, for example, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}. A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipient(s) (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}). Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}. Just as different habitats may favor different combinations of eco-morphological traits to maximize performance and fitness \citep{Arnold1983}, they may also shape different forms of a signal, so as to maximize its transmission and detection (e.g. \citealt{Seehausen1997}), or reduce its detection by unintended recipients such as predators \citep{Endler1984,Endler1990,Endler1991,Halfwerk2014}. This selective pressure may drive the local adaptation of communication signals.\\

@@

One potential barrier to the maintenance of localized signal divergence is the homogenizing effect of gene flow. Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}. This has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}. Yet, examples of microgeographic adaptation, i.e. adaptation at smaller scales than the range of dispersal, exist, highlighting a high potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein). Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, or local adaptation to predation pressure in North American salamanders \citep{Richardson2013}. Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.\\

@@

Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}. A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015}, territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}. Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}. Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\

Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002a, Thorpe2002b, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}. Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}. Other studies testing this hypothesis, however, found no pattern \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\

% So what? Introducing Anolis sagrei as a model species

% I am re-citing the same papers here, is that okay?
% This is the room for sagrei in particular

@@

Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km). Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}.

>>>>

The staggering diversity of animal communication signals has long been of interest to evolutionary biologists. Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}. A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipients (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}). Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}. Just as different habitats may favor different combinations of eco-morphological traits to maximize performance and fitness \citep{Arnold1983}, they may also shape different forms of a signal, so as to maximize its transmission and detection (e.g. \citealt{Seehausen1997}), or reduce its detection by unintended recipients such as predators \citep{Endler1984,Endler1990,Endler1991,Halfwerk2014}. This selective pressure may drive the local adaptation of communication signals.\\

@@

One potential barrier to the maintenance of localized signal divergence is the homogenizing effect of gene flow. Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}. This genetic homogenization has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}. Yet, examples of microgeographic adaptation, i.e. adaptation at smaller scales than the range of dispersal, exist, highlighting the potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein). Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, or local adaptation to predation pressure in North American salamanders \citep{Richardson2013}. Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.\\

@@

Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}. A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}. Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}. Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\

Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002a, Thorpe2002b, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}. Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}. Although this claim is further supported by recent findings that dewlap colors are perceived differently under different levels of shading \citep{Fleishman2020}, other studies found conflicting patterns of between-habitat variation that did not support the sensory drive hypothesis \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\

@@

Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km. Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}.
